\LARGE \bf Multi-Arm Bandit Models for 2D Grasp Planning with Uncertainty [v7 Feb 18, 2015 ]

### 2015-03-11 paper.tex
got figures

    \caption{\todo{decide on a name for Ben's method and try a variance approach as well}Comparison of the current average probability of force closure vs. the stopping time $T_s$.Graph is averaged over 100 shapes randomly drawn from the Brown Vision 2D Lab Dataset \cite{brown} with a set $|G|=1000$ for each shape.  We demonstrate this for Thompson, Gittins, Monte-Carlo and a prior Adaptive Sampling approach \cite{kehoe2012toward}. We also demonstrate what the average probability of force closure for the approximate optimal policy. Empirically, it appears that Thompson and Gittins converge at a faster rate to the optimal solution, which is desired for an anytime algorithm  }%
    \label{fig:motivation}%
\end{figure}

@@

Recent work has studied the effects of uncertainty in object pose and gripper positioning.%, motivated by part tolerancing from manufacturing and the use of low-cost sensors such as the Kinect.
Brook, Ciocarlie, and Hsiao~\cite{brook2011collaborative, hsiao2011bayesian} studied a Bayesian framework to evaluate the probability of grasp success given uncertainty in object identity, gripper positioning, and pose by simulating grasps on deterministic mesh and point cloud models.
Weisz et al.~\cite{weisz2012pose} found that grasps ranked by probability of force closure subject to uncertainty in object pose were empirically more successful on a physical robot grasps planned using deterministic wrench space metrics on objects from the Columbia Grasp Database.
Kim et al.~\cite{kim2012physically} planned grasps using dynamic simulations over perturbations in object pose, and also found that the planned grasps were more successful on a physical robot than classical grasp metrics.

@@

Some probabilistic grasp quality measures, such as probabilty of force closure, use Monte-Carlo sampling to evaluate grasp quality~\cite{christopoulos2007handling, kehoe2012estimating, kehoe2012toward}.
This involves sampling from distributions on random quantities and averaging the quality over these samples to empirically estimate a probability distribution~\cite{caflisch1998monte}.
It can be computationally expensive though to sample all proposed grasps to convergence.
To address the computational cost, Kehoe et al.~\cite{kehoe2012estimating} proposed an adaptive sampling procedure known as Iterative Pruning, which at a set iteration discards a subset of the grasps that seem unlikely to be of high quality . However, the method pruned grasps using only the sample mean and did not utilize variance around the current estimate, which can discard good grasps.
We propose modeling the problem as a Multi-Armed Bandit, which selects the next grasp to sample based on past observations instead ~\cite{barto1998reinforcement, lai1985asymptotically}.

@@

At each time step the MAB algorithm incurs \textit{regret}, the difference between the expected reward of the best arm and that of the arm selected. Bandit algorithms minimize cumulative regret, the sum of regret over the entire sequence of arm choices.
Lai and Robbins showed that the cumulative regret of the optimal solution to the bandit problem is bounded by a logarithmic function of the number of arm pulls~\cite{lai1985asymptotically}.
They presented an algorithm called (Upper Confidence Bound) UCB that obtains this bound asymptotically~\cite{lai1985asymptotically}.
The algorithm maintains a confidence bound on the distribution of reward based on prior observations and pulls the arm with the highest upper confidence bound.
%Many variants of UCB have thus been proposed~\cite{cesa2006prediction}. 
Since then several other algorithms have been shown to achieve near this bound in terms of convergence, such as the Gittins index policy~\cite{weber1992gittins} and Thompson sampling for the two arm case\cite{agrawal2011analysis}. \todo{The last result is actually only a few years ago. I can talk about contextual bandits for the sake of other recent information, but we don't use them and we are short on space. So I am wondering your opinion here}

@@

\subsubsection{The Gittins Index Method} 
One MAB method is to treat the problem as an Markov Decision Process (MDP) and use Markov Decision theory.
Formally, a MDP is defined as a set of possible of states, a set of actions, a set of transition probabilities between states, a reward function, and a discount factor \cite{barto1998reinforcement}.
In the Beta-Bernoulli MAB case, the set of actions is the $K$ options and the states are the Beta posterior on each option, or the integer values of the $\alpha$ and $\beta$ parameters.

Methods such as Value Iteration can compute optimal policies for an MDP with respect to the discount factor $\gamma$ when all states, actions, and expected rewards can be enumerated~\cite{weber1992gittins,barto1998reinforcement}.
However, the curse of dimensionality effects performance because if you have $K$ arms, a finite horizon of $T$ and a Beta-Bernoulli distribution on your options then your state space is exponential in $K$.
A key insight was given by Gittins, who showed that instead of solving the $K$-dimensional MDP one can instead solve $K$ 1-dimensional optimization problems: for each option $i$, $i= 1,...,k$, and for each state $x_t^i = \lbrace \alpha_0 +S_t, \beta_0 +F_t \rbrace^i$, where $S_t$ and $F_t$ correspond to the number of success and failures at pull $t$ and the state $x_t^i$ is the Beta prior for option $i$, as shown in Algorithm 1.

@@

\subsubsection{Thompson Sampling}
Computation of the Gittins indices can increase exponentially as the discount factor approaches $1$, and therefore it is not ideal to use in most cases. 
Thompson sampling is a less computationally expensive alternative.
All arms are initialized with a prior Beta distributions, which is normally Beta($\alpha=1$,  $\beta =1$) to reflect a uniform prior on the $\theta$ of the Bernoulli distribution.
Then for each arm draw $\theta_{j,t} \sim \mbox{Beta}(\alpha,\beta)$ and pull the arm with the highest $\theta_{j,t}$ drawn.
The reward, $X_{i,t}$ is observed from that arm, $j$, and the corresponding Beta distribution is updated.
This is repeated until a stopping time is reached. The full algorithm is shown in Algorithm 2.

@@

Uncertainty in object shape results from sensor noise and missing sensor data, which can occur due to transparency, specularity, and occlusions~\cite{mahler2015gp}.
Following ~\cite{mahler2015gp}, we represent the distribution over possible surfaces given sensing noise using a Gaussian process implicit surface (GPIS).
A GPIS represents a distribution over signed distance functions (SDFs).
A SDF is a real-valued function $f: \mathbb{R}^d \rightarrow \mathbb{R}$ that is greater than 0 outside the object, 0 on the surface and less than 0 inside the object.
A GPIS is a gaussian distribution over SDF values at a fixed set of query points $\mX = \{\bx_1, ... \bx_n\}, \bx_i \in \mathbb{R}^d$, $f(\bx_i) \sim \mN(\mu_{f}(\bx_i),\Sigma_{f}(\bx_i))$, where $\mu_{f}(\cdot)$ and $\Sigma_{f}(\cdot)$ are the mean and covariance functions of the GPIS~\cite{rasmussen2006}.
See Mahler et al. for details on how to estimate a mean and covariance function and sample shapes from a GPIS~\cite{mahler2015gp}.
Following Mahler et al., we set $\mX$ to a uniform $M \times M$ grid of points with square cells.
For convenience, in later sections we will refer to the GPIS parameters as $\theta = \left( \mu_{f}(x), \Sigma_{f}(x) \right)$.

@@

 In Fig. \ref{fig:pulls_per_grasp}, we plot samples per grasp quality. Gittins and Thompson appear to allocate  more samples to grasps of high quality in contrast to uniform allocation and Iterative Pruning. The ability to find grasps with a higher probability of force  faster and have a smaller confidence interval on the returned quantity can be explain by the fact that MAB methods hone in their sampling efforts on promising grasps, for the top 10$\%$ of grasps Thompson Sampling allocates 29$\%$ of the total number of samples and Gittins allocates 23$\%$ of total samples. Uniform allocation and Iterative Pruning allocate $1\%$  and $6\%$ of total samples, respectively.

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.25]{matlab_figures/pfc.eps}
\caption{ \footnotesize Comparison of the current average probability of force closure vs. the stopping time $T_s$.Graph is averaged over 100 objects randomly drawn from the Brown Vision 2D Lab Dataset \cite{brown} with a set $|G|=1000$ for each object.  We demonstrate this for Thompson, Gittins, Monte-Carlo and the Adaptive Sampling approach \cite{kehoe2012toward}. We also demonstrate what the average probability of force closure for the approximate optimal policy. Empirically, it appears that Thompson and Gittins converge at a faster rate to the optimal solution, which is desired for an anytime algorithm  }
\vspace*{-10pt}
\label{fig:grasp_quality}
\end{figure}

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.25]{matlab_figures/pulls_per_grasp.eps}
\caption{ \footnotesize Comparison of sample per grasp for the four sequential decision methods (Monte-Carlo, Thompson, Gittins). Graph is averaged over 100 objects from the Brown Silhouette Dataset \cite{brown} with a set $|G|=1000$ for each object. The best grasps are ranked 1 and worst are 1000. As illustrated the MAB algorithms intelligently allocate samples towards high quality grasps based on past observations, where Monte-Carlo Integration takes a uniform approach to allocation. }

@@

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.25]{matlab_figures/confidence_bounds.eps}
\caption{ \footnotesize \todo{We might want to condense Fig 6 and Fig. 4. The confidence intervals clutter the plot though} The probability of Force Closure of the grasp returned vs the number of samples required with a $95\%$ confidence interval around the return value for Thompson Sampling, the adaptive sampling method and Monte-Carlo sampling. Graph is averaged over 100 objects from the Brown Silhouette Dataset \cite{brown} with a set $|G|=1000$ for each object. The sharper decrease in confidence interval corresponds to more samples being allocated for the return grasp. }

@@

    \caption{\todo{Need to change this figure, wondering what should be conveyed here}Three objects shown from the Brown Visual Lab Dataset with induced shape uncertainty and visualized with the GPIS-Blur method \cite{mahler2015gp}. The four methods (Monte-Carlo, Kehoe's, Thompson and Gittins) were all run until a stopping time of $10,000$ evaluations with a uniformly initialized grasp set of $|G|=1000$. We also showed the estimated best grasp in the set.  The grasps and the quality each one found is shown above.  In the two out of three cases, Thompson sampling is able to find the best grasp in the set at the stopping interval of $10,000$, however at the last shape there is a difference of $2\%$ grasp quality.   }%
    \label{fig:shape_samples}%
\end{figure*}

@@

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.25]{matlab_figures/worst_case.eps}
\caption{ \footnotesize Comparison of the current average probability of force closure vs. the stopping time $T_s$ for the pathological case where the samples are sorted to be misleading. The graph is averaged over 100 objects randomly drawn from the Brown Vision 2D Lab Dataset \cite{brown} with a set $|G|=1000$ for each object.  We demonstrate this for Thompson, Gittins, Monte-Carlo and the approach taking in Kehoe et al \cite{kehoe2012toward}. We also demonstrate what the average probability of force closure for the approximate optimal policy. Empirically, it appears that when samples are misleading the best policy is uniform allocation. However, it also illustrated that Thompson sampling can recover from this situation where Gittins and Kehoe et al.  are not able to and get stuck in local solutions. }
\vspace*{-10pt}
\label{fig:worst_case}
\end{figure}

>>>>

    \caption{\todo{decide on a name for Ben's method and try a variance approach as well}Comparison of the current average probability of force closure vs. the stopping time $T_s$.Graph is averaged over 100 shapes randomly drawn from the Brown Vision 2D Lab Dataset \cite{brown} with a set $|G|=1000$ for each shape.  We demonstrate this for Thompson, Gittins, uniform allocation and Iterative Pruning \cite{kehoe2012estimating}. We also demonstrate what the average highest probability of force closure is when the brute force approach of exhaustively sampling each candidate grasp is used. Empirically, it appears that Thompson and Gittins converge at a faster rate to the optimal solution, which is desired for an anytime algorithm  }%
    \label{fig:motivation}%
\end{figure}

@@

Recent work has studied the effects of uncertainty in object pose and gripper positioning.
 Brook, Ciocarlie, and Hsiao~\cite{brook2011collaborative, hsiao2011bayesian} studied a Bayesian framework to evaluate the probability of grasp success given uncertainty in object identity, gripper positioning, and pose by simulating grasps on deterministic mesh and point cloud models.
Weisz et al.~\cite{weisz2012pose} found that grasps ranked by probability of force closure subject to uncertainty in object pose were empirically more successful on a physical robot grasps planned using deterministic wrench space metrics on objects from the Columbia Grasp Database.
Kim et al.~\cite{kim2012physically} planned grasps using dynamic simulations over perturbations in object pose, and also found that the planned grasps were more successful on a physical robot than classical grasp metrics.

@@

Some probabilistic grasp quality measures, such as probabilty of force closure, use Monte-Carlo sampling to evaluate grasp quality~\cite{christopoulos2007handling, kehoe2012estimating, kehoe2012toward}.
This involves sampling from distributions on random quantities and averaging the quality over these samples to empirically estimate a probability distribution~\cite{caflisch1998monte}.
It can be computationally expensive though to sample all proposed grasps to convergence.
To address the computational cost, Kehoe et al.~\cite{kehoe2012estimating} proposed an adaptive sampling procedure known as Iterative Pruning, which at a set iteration discards a subset of the grasps that seem unlikely to be of high quality . However, the method pruned grasps using only the sample mean and did not utilize any confidence interval around the current estimate, which can discard good grasps.
We propose modeling the problem as a Multi-Armed Bandit, which selects the next grasp to sample based on past observations instead ~\cite{barto1998reinforcement, lai1985asymptotically}.

@@

At each time step the MAB algorithm incurs \textit{regret}, the difference between the expected reward of the best arm and that of the arm selected. Bandit algorithms minimize cumulative regret, the sum of regret over the entire sequence of arm choices.
Lai and Robbins showed that the cumulative regret of the optimal solution to the bandit problem is bounded by a logarithmic function of the number of arm pulls~\cite{lai1985asymptotically}.
They presented an algorithm called (Upper Confidence Bound) UCB that obtains this bound asymptotically~\cite{lai1985asymptotically}.
The algorithm maintains a confidence bound on the distribution of reward based on prior observations and pulls the arm with the highest upper confidence bound.
%Many variants of UCB have thus been proposed~\cite{cesa2006prediction}. 
Since then several other algorithms have been shown to achieve near this bound in terms of convergence, such as the Gittins index policy~\cite{weber1992gittins} and Thompson sampling for the 2 arm case\cite{agrawal2011analysis}. \todo{The last result is actually only a few years ago. I can talk about contextual bandits for the sake of other recent information, but we don't use them and we are short on space. So I am wondering your opinion here}

@@

\subsubsection{The Gittins Index Method} 
One MAB method is to treat the problem as a Markov Decision Process (MDP) and use Markov Decision theory.
Formally, a MDP is defined as a set of possible of states, a set of actions, a set of transition probabilities between states, a reward function, and a discount factor \cite{barto1998reinforcement}.
In the Beta-Bernoulli MAB case, the set of actions is the $K$ options and the states are the Beta posterior on each option, or the integer values of the $\alpha$ and $\beta$ parameters.

Methods such as Value Iteration can compute optimal policies for an MDP with respect to the discount factor $\gamma$ when all states, actions, and expected rewards can be enumerated~\cite{weber1992gittins,barto1998reinforcement}.
However, the curse of dimensionality effects performance because for $K$ arms, a finite horizon of $T$ and a Beta-Bernoulli distribution on each arm then the state space is exponential in $K$.
A key insight was given by Gittins, who showed that instead of solving the $K$-dimensional MDP one can instead solve $K$ 1-dimensional optimization problems: for each option $i$, $i= 1,...,k$, and for each state $x_t^i = \lbrace \alpha_0 +S_t, \beta_0 +F_t \rbrace^i$, where $S_t$ and $F_t$ correspond to the number of success and failures at pull $t$ and the state $x_t^i$ is the Beta prior for option $i$, as shown in Algorithm 1.

@@

\subsubsection{Thompson Sampling}
Computation of the Gittins indices can increase exponentially in time as the discount factor approaches $1$, and therefore it is not ideal to use in most cases. 
Thompson sampling is a less computationally expensive alternative.
All arms are initialized with a prior Beta distributions, which is normally Beta($\alpha=1$,  $\beta =1$) to reflect a uniform prior on the $\theta$ of the Bernoulli distribution.
Then for each arm draw $\theta_{j,t} \sim \mbox{Beta}(\alpha,\beta)$ and pull the arm with the highest $\theta_{j,t}$ drawn.
The reward, $X_{i,t}$ is observed from that arm, $j$, and the corresponding Beta distribution is updated.
This is repeated until a stopping time is reached. The full algorithm is shown in Algorithm 2.

@@

Uncertainty in object shape results from sensor noise and missing sensor data, which can occur due to transparency, specularity, and occlusions~\cite{mahler2015gp}.
Following ~\cite{mahler2015gp}, we represent the distribution over possible surfaces given sensing noise using a Gaussian process implicit surface (GPIS).
A GPIS represents a distribution over signed distance functions (SDFs).
A SDF is a real-valued function $f: \mathbb{R}^d \rightarrow \mathbb{R}$ that is greater than 0 outside the object, 0 on the surface and less than 0 inside the object.
A GPIS is a Gaussian distribution over SDF values at a fixed set of query points $\mX = \{\bx_1, ... \bx_n\}, \bx_i \in \mathbb{R}^d$, $f(\bx_i) \sim \mN(\mu_{f}(\bx_i),\Sigma_{f}(\bx_i))$, where $\mu_{f}(\cdot)$ and $\Sigma_{f}(\cdot)$ are the mean and covariance functions of the GPIS~\cite{rasmussen2006}.
See Mahler et al. for details on how to estimate a mean and covariance function and sample shapes from a GPIS~\cite{mahler2015gp}.
Following Mahler et al., we set $\mX$ to a uniform $M \times M$ grid of points with square cells.
For convenience, in later sections we will refer to the GPIS parameters as $\theta = \left( \mu_{f}(x), \Sigma_{f}(x) \right)$.

@@

 In Fig. \ref{fig:pulls_per_grasp}, we plot samples per grasp quality. Gittins and Thompson appear to allocate  more samples to grasps of high quality in contrast to uniform allocation and Iterative Pruning. The ability to find grasps with a higher probability of force  faster and have a smaller confidence interval on the returned quantity can be explain by the fact that in our experiments the MAB methods hone in their sampling efforts on promising grasps, for the top 10$\%$ of grasps Thompson Sampling allocates 29$\%$ of the total number of samples and Gittins allocates 23$\%$ of total samples. Uniform allocation and Iterative Pruning allocate $1\%$  and $6\%$ of total samples, respectively.

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.25]{matlab_figures/pfc.eps}
\caption{ \footnotesize Comparison of the current average probability of force closure vs. the stopping time $T_s$.Graph is averaged over 100 objects randomly drawn from the Brown Vision 2D Lab Dataset \cite{brown} with a set $|G|=1000$ for each object.We demonstrate this for Thompson, Gittins, uniform allocation and Iterative Pruning \cite{kehoe2012estimating}. We also demonstrate what the average highest probability of force closure is when the brute force approach of exhaustively sampling each candidate grasp is used. Empirically, it appears that Thompson and Gittins converge at a faster rate to the optimal solution, which is desired for an anytime algorithm  }
\vspace*{-10pt}
\label{fig:grasp_quality}
\end{figure}

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.25]{matlab_figures/pulls_per_grasp.eps}
\caption{ \footnotesize Comparison of sample per grasp for the four sequential decision methods (Thompson, Gittins, uniform allocation and Iterative Pruning). Graph is averaged over 100 objects from the Brown Silhouette Dataset \cite{brown} with a set $|G|=1000$ for each object. The best grasps are ranked 1 and worst are 1000. As illustrated the MAB algorithms intelligently allocate samples towards high quality grasps based on past observations, where Monte-Carlo Integration takes a uniform approach to allocation. }

@@

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.25]{matlab_figures/confidence_bounds.eps}
\caption{ \footnotesize \todo{We might want to condense Fig 6 and Fig. 4. The confidence intervals clutter the plot though} The probability of Force Closure of the grasp returned vs the number of samples required with a $95\%$ confidence interval around the return value for Thompson Sampling, Iterative Pruning and uniform allocation. Graph is averaged over 100 objects from the Brown Silhouette Dataset \cite{brown} with a set $|G|=1000$ for each object. The sharper decrease in confidence interval corresponds to more samples being allocated for the return grasp. }

@@

    \caption{\todo{Need to change this figure, wondering what should be conveyed here}Three objects shown from the Brown Visual Lab Dataset with induced shape uncertainty and visualized with the GPIS-Blur method \cite{mahler2015gp}. The four methods (uniform allocation, Iterative Pruning, Thompson and Gittins) were all run until a stopping time of $10,000$ evaluations with a randomly selected grasp set of $|G|=1000$. We also show the grasp returned by the brute force approach of exhaustively sampling every grasp in the set.  The grasps and the quality each one found is shown above.  In the two out of three cases, Thompson sampling is able to find the best grasp in the set at the stopping interval of $10,000$, however at the last shape there is a difference of $2\%$ grasp quality.   }%
    \label{fig:shape_samples}%
\end{figure*}

@@

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.25]{matlab_figures/worst_case.eps}
\caption{ \footnotesize Comparison of the current average probability of force closure vs. the stopping time $T_s$ for the pathological case where the samples are sorted to be misleading. The graph is averaged over 100 objects randomly drawn from the Brown Vision 2D Lab Dataset \cite{brown} with a set $|G|=1000$ for each object.  We demonstrate this for Thompson, Gittins, uniform allocation and Iterative Pruning \cite{kehoe2012estimating}. We also demonstrate what the average highest probability of force closure is when the brute force approach of exhaustively sampling each candidate grasp is used. Empirically, it appears that when samples are misleading the best policy is uniform allocation. However, it also illustrated that Thompson sampling can recover from this situation where Gittins and Kehoe et al.  are not able to and get stuck in local solutions. }
\vspace*{-10pt}
\label{fig:worst_case}
\end{figure}
